MSTK: Mesh Toolkit, v 1.7 - DRAFT

### 2010-02-25 docs/MSTK.tex
Updated!
Added section on volatile entities and description for related routines

{\bf NOTE: CURRENTLY THE F1 REPRESENTATION IS THE MOST ROBUST
  IMPLEMENTATION SINCE IT IS MOST COMMONLY USED. THIS REPRESENTATION
  HAS GENERALLY BEEN STABLE FOR SEVERAL YEARS. THE REDUCED
  REPRESENTATION HAVE NOT BEEN AS HEAVILY DEBUGGED AND IT IS CURRENTLY
  BROKEN. IF YOU CHOOSE TO USE SOME OF THE OTHER REPRESENTATIONS
  PLEASE REPORT ANY PROBLEMS IMMEDIATELY TO \texttt{rao@lanl.gov}.  }

@@

@@

\item[]{\bf {\em void} MESH\_Rem\_Region({\em Mesh\_ptr}
    mesh, {\em MRegion\_ptr} r):} Remove region from mesh. Region
  is not deleted and must be deleted afterward separately. Normally,
  one need not call this since {\bf {\em MR\_Delete}} will
  remove the region from the mesh. Use this only if you know exactly
  what you are doing.
\end{description}

@@

@@

\item[]{\bf {\em void} MF\_Coords({\em MFace\_ptr} mface,
{\em int} *n, double (*xyz)[3], {\em int} dir):} Returns the
coordinates of the face vertices in an array along with the number of
vertices. 
% If 'dir' is 1, the coordinates are returned with a  ccw
% ordering while looking down the face normal; if 'dir' is 0, they are
% returned with a cw ordering (See Figure~\ref{fig:face_def}).
\end{description}

>>>>

{\bf NOTE: CURRENTLY THE F1 REPRESENTATION IS THE MOST ROBUST
  IMPLEMENTATION SINCE IT IS MOST COMMONLY USED. THIS REPRESENTATION
  HAS GENERALLY BEEN STABLE FOR SEVERAL YEARS. THE REDUCED
  REPRESENTATIONS HAVE NOT BEEN AS HEAVILY DEBUGGED AND IT MAY HAVE
  SOME BUGS. IF YOU CHOOSE TO USE SOME OF THE OTHER REPRESENTATIONS
  PLEASE REPORT ANY PROBLEMS IMMEDIATELY TO \texttt{rao@lanl.gov}.  }

@@

\subsection{Volatile mesh entities in reduced representations}

As discussed above reduced representations, do not explicitly
represent all types of mesh entities. For example, R1 representations
do not explicitly represent edges and faces while R4 representations
do explicitly represent edges only. However, whenever an implicit
entity is requested from MSTK, the software creates a temporary entity
so that it appears that the entity actually exists in the
database. Since these entities are created-on-the-fly they are called
{\em volatile} mesh entities. For example, if an application asked for
the faces of a region in an R1 representation, MSTK will create as
many volatile faces as necessary, put them in a list and return them
to the calling application. Thus the application can pretend to work
with a full representation (although it is not always efficient to do so).

Volatile entities are stored for a period of time in internal data
structures in MSTK. Whenever there is a request for a new volatile
mesh entity, the code first checks if this entity has already been
created. If the volatile already exists in the database, then the
entity is returned as is and if it does not, it is created. This
ensures that many copies of the same entity are not created in
MSTK. However, to ensure that MSTK also does not store every volatile
entity forever (thereby, recreating a full representation), there is a
mechanism to perform garbage cleaning on volatile
entities. Periodically through the execution of the code, volatile
entities that have not been used for a long time are deleted from the
database freeing up valuable memory (which was the main point of using
a reduced representation after all).

To ensure that garbage cleaning does not delete a volatile mesh entity
that is being processed or stored by a calling application, MSTK also
provides a mechanism for locking entities. One can specify an autolock
mechanism for the entire mesh which means that no volatile entity will
ever be deleted once it is created. On the other hand, applications
may lock specific entities and unlock them when they cease to be
useful.

@@

\item[]{\bf {\em void} MESH\_Rem\_Region({\em Mesh\_ptr}
    mesh, {\em MRegion\_ptr} r):} Remove region from mesh. Region
  is not deleted and must be deleted afterward separately. Normally,
  one need not call this since {\bf {\em MR\_Delete}} will
  remove the region from the mesh. Use this only if you know exactly
  what you are doing.

\item[]{\bf {\em void} MESH\_Set\_AutoLock({\em Mesh\_ptr} mesh, {\em
      int} autolock):} If autolock is 1, all volatile mesh entities
  created in reduced representations will be automatically locked and
  cannot be removed by garbage cleaning procedures (They can still be
  removed by explicitly calling a delete on them). If it 0, volatile
  mesh entities can be deleted internally after a period of disuse.

\item[]{\bf {\em int} MESH\_AutoLock({\em Mesh\_ptr} mesh):} Return
the autolock status for volatile entities (See {\bf
  MESH\_Set\_AutoLock} for details).

@@

\item[]{\bf {\em void} ME\_Lock({\em MEdge\_ptr} e):} Lock a volatile
  edge so that it cannot be deleted by garbage cleaning procedures (It
  can still be deleted by {\bf ME\_Delete}.

\item[]{\bf {\em void} ME\_UnLock({\em MEdge\_ptr} e):} Unlock a volatile
  edge so that it may be deleted freely by garbage cleaning procedures.

@@

\item[]{\bf {\em void} MF\_Coords({\em MFace\_ptr} mface,
{\em int} *n, double (*xyz)[3], {\em int} dir):} Returns the
coordinates of the face vertices in an array along with the number of
vertices. 
% If 'dir' is 1, the coordinates are returned with a  ccw
% ordering while looking down the face normal; if 'dir' is 0, they are
% returned with a cw ordering (See Figure~\ref{fig:face_def}).

\item[]{\bf {\em void} MF\_Lock({\em MFace\_ptr} f):} Lock a volatile
  face so that it cannot be deleted by garbage cleaning procedures (It
  can still be deleted by {\bf MF\_Delete}.

\item[]{\bf {\em void} MF\_UnLock({\em MFace\_ptr} f):} Unlock a volatile
  face so that it may be deleted freely by garbage cleaning procedures.
